---
title: "p8105_hw6_jt3174"
author: "Jingyi Tang"
date: "11/23/2019"
output: github_document
---

```{r set up}

# ensure reproductivity
set.seed(1)

# load library
library(tidyverse)
library(modelr)
library(mgcv)
library(ggplot2)
library(viridis)
library(ggridges)

knitr::opts_chunk$set(
  # display the code in the code truck above its results in the final document
  echo = TRUE,
  # do not display any warning messages generated by the code
  warning = FALSE,
  # set the figure to be 8 x 6, and the proportion it takes to be 90%
  fig.width = 8,
  fig.height = 6, 
  out.width = "90%"
)

# setting a global options for continuous data color family and a different format to set discrete data to have a color family
options(
  ggplot2.countinuous.colour = "viridis",
  ggplot2.countinuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# have a minimal theme and legends at the bottom
theme_set(theme_minimal() + theme(legend.position = "bottom"))

```


## Problem 1

```{r load and clean data}

# Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

baby = read_csv("./data/birthweight.csv") %>% 
  # clean names
  janitor::clean_names() %>% 
  # use case_when to give the values of certain variables new values
  mutate(
    # baby’s sex (male = 1, female = 2)
    babysex = as.factor(case_when(
      babysex == 1 ~ "male",
      babysex == 2 ~ "female"
    )),
    # father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
    frace = as.factor(case_when(
      frace == 1 ~ "White",
      frace == 2 ~ "Black",
      frace == 3 ~ "Asian",
      frace == 4 ~ "Puerto Rican",
      frace == 8 ~ "Other",
      frace == 9 ~ "Unknown"
    )),
    # presence of malformations that could affect weight (0 = absent, 1 = present)
    malform = as.factor(case_when(
      malform == 0 ~ "absent",
      malform == 1 ~ "present"
    )),
    # mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
    mrace = as.factor(case_when(
      mrace == 1 ~ "White",
      mrace == 2 ~ "Black",
      mrace == 3 ~ "Asian",
      mrace == 4 ~ "Puerto Rican",
      mrace == 8 ~ "Other"
    ))  
  )

# checking if there is any missing value in the fataframe called baby
# if all of the values in the table returned were 0's, there are no missing values
sapply(baby, function(x) sum(is.na(x))) %>% 
  knitr::kable(., col.names = "Number of Missing Value", caption = "Table for Variables and Their Missing Values")

```

There is no missing value in the dataset `baby`, and this dataset is cleaned as stated in the homework page.

For a regression model to propose, I decided to use anova to test nested model. I started with full regression, and then decided by the p-value of each factor in the model to decide to keep or not for the proposed model. And then I will use anova test these two models to decided the final proposed model in this step. 

```{r regression models}

# full regression model
full_model = lm(bwt ~ ., data = baby)

# diaplay the model summary
summary(full_model)$coefficient%>%
  knitr::kable(., col.names = c("Estimate", "Std. Error", "t value","Pr(>|t|)"), caption = "result of main effect linear regression model")

```

Here, accoding to the p-values, I am going to keep the varaibles `babysex`, `bhead`, `blength`, `delwt`, `gaweeks`, `parity`, `smoke`. 

```{r regression model proposed}

# proposed model with 
proposed_model = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + parity + smoken, data = baby)

# diaplay the model summary
summary(proposed_model)$coefficient%>%
  knitr::kable(., col.names = c("Estimate", "Std. Error", "t value","Pr(>|t|)"), caption = "result of main effect linear regression model")

# use anova() to decide which model is better
anova(proposed_model, full_model) %>% 
  broom::tidy()

```

We can see that the proposed model is a nested model for full model, and accoding to the F-test, p-value = 1.03464e-56 < significance level at 0.05, the proposed model is better than the full model. 

```{r regression model}
# add residual and prediction to the orginal dataframe
baby_residual_prediction = baby %>% 
  add_predictions(proposed_model) %>% 
  add_residuals(proposed_model)

# create residual vs fitted plot
baby_residual_prediction %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = .5, color = "darksalmon") +
  labs(title = "Residuals vs. Fitted Value Plot for Proposed Model",
       x = "Fitted Values",
       y = "Residuals")

```

Here, we fit two regression models stated in the homework page. 

```{r regression models to be compared for birthweight}

# One using length at birth and gestational age as predictors (main effects only)

main_effect_model = lm(bwt ~ blength + gaweeks, data = baby)

# diaplay the model summary
summary(main_effect_model)$coefficient%>%
  knitr::kable(., col.names = c("Estimate", "Std. Error", "t value","Pr(>|t|)"), caption = "result of main effect linear regression model")

# One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

interaction_model = lm(bwt ~ bhead * blength * babysex , data = baby)

# display the model summary
summary(interaction_model)$coefficient%>%
  knitr::kable(., col.names = c("Estimate", "Std. Error", "t value","Pr(>|t|)"), caption = "result of interaction linear regression model")

```

